---
title: "DA5020.P2.MirSultan.AsmaDjourdikh"
author: "Mir Sultan"
date: "2023-07-16"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Step 1: Identify the dimensions from each dimension table  
We have two dimension tables represented by business.csv and bed_type.csv.
The dimensions in bed_type.csv are are the bed codes, bed descriptions, and bed IDs. The attributes in this table or csv file are for the beds in the fact table bed_fact.csv.

The dimensions in business_type.csv are business name, IMS org ID, and bed cluster ID. 


# Step 2: Identify the Facts variables from the single Fact Table 
The facts in the fact table bed_fact.csv are license_beds, census_beds, and staffed_beds. 


# Step 3
```{r message=FALSE, warning=FALSE}
library(RSQLite)
library(DBI)
library(XML)
library(xml2)
library(RCurl)
library(dplyr)
library(stringr)
library(tidyverse)
library(ggplot2)
library(tidyr)
library(lubridate)
library(scales)
library(readr)

mydb <- dbConnect(RSQLite::SQLite())

```

```{sql, connection=mydb}
CREATE TABLE IF NOT EXISTS bed_type (
  bed_id INTEGER PRIMARY KEY,
  bed_code TEXT,
  bed_desc TEXT
);
```
```{sql, connection=mydb}
CREATE TABLE IF NOT EXISTS business (
  ims_org_id TEXT PRIMARY KEY,
  business_name TEXT,
  ttl_license_beds INTEGER,
  ttl_census_beds INTEGER,
  ttl_staffed_beds INTEGER,
  bed_cluster_id INTEGER,
  FOREIGN KEY(bed_cluster_id) REFERENCES bed_type(bed_id)
);
```
```{sql, connection=mydb}
CREATE TABLE IF NOT EXISTS bed_fact (
  ims_org_id TEXT,
  bed_id INTEGER REFERENCES bed_type(bed_id),
  license_beds INTEGER,
  census_beds INTEGER,
  staffed_beds INTEGER,
  FOREIGN KEY(ims_org_id) REFERENCES business(ims_org_id),
  FOREIGN KEY(bed_id) REFERENCES bed_type(bed_id)
);
```

```{r}
bed_type <- read.csv("bed_type.csv", stringsAsFactors = FALSE)
bed_fact <- read.csv("bed_fact.csv", stringsAsFactors = FALSE)
business <- read.csv("business.csv", stringsAsFactors = FALSE)

dbWriteTable(mydb, "bed_type", bed_type, append = TRUE)
dbWriteTable(mydb, "bed_fact", bed_fact, append = TRUE)
dbWriteTable(mydb, "business", business, append = TRUE)

```
For better SQL practice, everything in the CSV files were made into SQL tables where the appropriate primary and foreign keys were applied.

## Step 3a: Analysis for Leadership 
```{sql, connection=mydb}
SELECT b.business_name AS hospital_name, SUM(bf.license_beds) AS total_license_beds
FROM business b
JOIN bed_fact bf ON b.ims_org_id = bf.ims_org_id
WHERE bf.bed_id IN (4, 15)
GROUP BY b.business_name
ORDER BY total_license_beds DESC
LIMIT 10;
```

```{sql, connection=mydb}
SELECT b.business_name AS hospital_name, SUM(bf.census_beds) AS total_census_beds
FROM business b
JOIN bed_fact bf ON b.ims_org_id = bf.ims_org_id
WHERE bf.bed_id IN (4, 15)
GROUP BY b.business_name
ORDER BY total_census_beds DESC
LIMIT 10;
```

```{sql, connection=mydb}
SELECT b.business_name AS hospital_name, SUM(bf.staffed_beds) AS total_staffed_beds
FROM business b
JOIN bed_fact bf ON b.ims_org_id = bf.ims_org_id
WHERE bf.bed_id IN (4, 15)
GROUP BY b.business_name
ORDER BY total_staffed_beds DESC
LIMIT 10;
```

## Step 3b: Interpretation of findings 

From the top Hospitals by License Beds (table 1) we found:
Phoenix Children's Hospital has the highest number of license beds with 247. University of Maryland Medical Center and UC Health University Hospital follow closely with 220 and 218 license beds, respectively. Wesley Medical Center, LLC and Vidant Medical Center also have significant bed volumes with 214 and 204 license beds, respectively.

From the top Hospitals by Census Beds (table 2) we found:
Shands Hospital at the University of Florida has the highest number of census beds with 167. Dallas County Hospital Association and Mercy Medical Center Saint Louis are also prominent with 145 and 142 census beds, respectively. The Methodist Hospital and University of Minnesota Medical Center Fairview are among the top hospitals as well.

From the top Hospitals by Staffed Beds (table 3) we found:
Vidant Medical Center and Rady Childrens Hospital and Health Center both have the highest number of staffed beds with 203 and 200, respectively. University of Maryland Medical Center, Emory University Hospital, and Shands Hospital at the University of Florida are also notable with significant staffed bed volumes.

After we identify the top hospitals based on licensed beds, staffed beds, and Census beds, we noticed that the following hospitals are appearing on multiple table, Vidant Medical Center appears in the top hospitals for both license beds and staffed beds, indicating its prominence in bed volume. University of Maryland Medical Center is present in the top hospitals for license beds and staffed beds, suggesting its significance in both bed availability and utilization. Wesley Medical Center, LLC appears in the top hospitals for both license beds and staffed beds, making it another potential candidate for the intervention pilot program.
Based on these findings, Phoenix Childrens Hospital, UC Health University Hospital, and Rady Childrens Hospital and Health Center are notable hospitals with high bed volumes. These hospitals, along with the ones appearing on multiple lists, can be considered as potential candidates for the intervention pilot program.

# Step 4
```{r warning=FALSE}
"Query Process Steps
1. Getting Data (From, Join)
2. Row Filter (Where)
3. Grouping (Group by)
4. Group Filter (Having)
5. Return Expressions (Select)
6. Order & Paging (Order by & Limit / Offset)"
```

## Step 4a: Drill down investigation 
```{sql, connection=mydb}
SELECT b.business_name AS hospital_name, SUM(bf.license_beds) AS total_license_beds
FROM business b
JOIN bed_fact bf ON b.ims_org_id = bf.ims_org_id
WHERE bf.bed_id IN (4, 15)
GROUP BY b.business_name
HAVING COUNT(DISTINCT bf.bed_id) = 2
ORDER BY total_license_beds DESC
LIMIT 10;
```

```{sql, connection=mydb}
SELECT b.business_name AS hospital_name, SUM(bf.census_beds) AS total_census_beds
FROM business b
JOIN bed_fact bf ON b.ims_org_id = bf.ims_org_id
WHERE bf.bed_id IN (4, 15)
GROUP BY b.business_name
HAVING COUNT(DISTINCT bf.bed_id) = 2
ORDER BY total_census_beds DESC
LIMIT 10;
```

```{sql, connection=mydb}
SELECT b.business_name AS hospital_name, SUM(bf.staffed_beds) AS total_staffed_beds
FROM business b
JOIN bed_fact bf ON b.ims_org_id = bf.ims_org_id
WHERE bf.bed_id IN (4, 15)
GROUP BY b.business_name
HAVING COUNT(DISTINCT bf.bed_id) = 2
ORDER BY total_staffed_beds DESC
LIMIT 10;
```

##Interpretation of findings

Top Hospitals by License Beds (having both bed IDs 4 and 15):
University of Maryland Medical Center has the highest number of license beds with 220.
UC Health University Hospital follows closely with 218 license beds.
Shands Hospital at the University of Florida is also prominent with 167 license beds.
MCGHealth, Inc and Grady Memorial Hospital are among the top hospitals in terms of license bed volume.

Top Hospitals by Census Beds (having both bed IDs 4 and 15):
Shands Hospital at the University of Florida has the highest number of census beds with 167.
University of Minnesota Medical Center Fairview follows with 129 census beds.
University of Maryland Medical Center is also notable with 127 census beds.
Jackson Memorial Hospital and UC Health University Hospital are among the top hospitals in terms of census bed volume.

Top Hospitals by Staffed Beds (having both bed IDs 4 and 15):
University of Maryland Medical Center has the highest number of staffed beds with 171.
Shands Hospital at the University of Florida follows closely with 167 staffed beds.
Grady Memorial Hospital and UC Health University Hospital are also notable with significant staffed bed volumes.
University of Minnesota Medical Center Fairview and Carolinas Medical Center are among the top hospitals in terms of staffed bed volume.

Also, in this step we can see the hospitals Appearing on Multiple Lists(having both bed IDs 4 and 15). These hospitals are; University of Maryland Medical Center that appears in the top hospitals for license beds, census beds, and staffed beds, suggesting its prominence in bed availability, utilization, and staffing. UC Health University Hospital appears in the top hospitals for license beds, census beds, and staffed beds, making it another potential candidate for the intervention pilot program. Shands Hospital at the University of Florida is present in the top hospitals for both census beds and staffed beds, indicating its significance in both bed utilization and staffing. Grady Memorial Hospital and University of Minnesota Medical Center Fairview are also notable hospitals appearing in multiple lists.
Based on these findings, University of Maryland Medical Center, UC Health University Hospital, Shands Hospital at the University of Florida, Grady Memorial Hospital, and University of Minnesota Medical Center Fairview are hospitals that consistently appear in the top lists across different bed categories. These hospitals can be considered as strong candidates for the intervention pilot program due to their high bed volumes and presence in multiple lists.


##Step 4b: Final recommendation 

Based on the analysis conducted in steps 3a and 4a, my final recommendation for Leadership is to consider the following hospitals as the best candidates for their pilot intervention program:

University of Maryland Medical Center: This hospital consistently appears in the top hospitals across all three bed categories (license beds, census beds, and staffed beds). It has a high bed volume and exhibits strong performance in terms of bed availability, utilization, and staffing. Selecting University of Maryland Medical Center as a pilot site would provide Leadership with an opportunity to observe the impact of their intervention program in a hospital with a significant bed capacity and utilization.

UC Health University Hospital: Similar to University of Maryland Medical Center, UC Health University Hospital demonstrates consistent performance across all three bed categories. It has a notable bed volume and performs well in terms of both bed availability and staffed beds. Choosing UC Health University Hospital as a pilot site would allow Leadership to assess the effectiveness of their intervention program in a hospital with a strong bed infrastructure and staffing capabilities.

By focusing on these two hospitals, Leadership can concentrate their efforts on implementing the intervention program in settings that have demonstrated high bed volumes and appear consistently in multiple lists. This approach aligns with Leadership's goal of selecting one or two hospitals for their pilot sites while maximizing the potential impact of the intervention program.










