---
title: "DA5020.P2.Mir.Sultan"
author: "Mir Sultan"
date: "2023-07-16"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Step 1
We have two dimension tables represented by business.csv and bed_type.csv.
The dimensions in bed_type.csv are are the bed codes, bed descriptions, and bed IDs. The attributes in this table or csv file are for the beds in the fact table bed_fact.csv.

The dimensions in business_type.csv are business name, IMS org ID, and bed cluster ID. 


# Step 2
The facts in the fact table bed_fact.csv are license_beds, census_beds, and staffed_beds. 


# Step 3
```{r message=FALSE, warning=FALSE}
library(RSQLite)
library(DBI)
library(XML)
library(xml2)
library(RCurl)
library(dplyr)
library(stringr)
library(tidyverse)
library(ggplot2)
library(tidyr)
library(lubridate)
library(scales)
library(readr)

mydb <- dbConnect(RSQLite::SQLite())

```

```{sql, connection=mydb}
CREATE TABLE IF NOT EXISTS bed_type (
  bed_id INTEGER PRIMARY KEY,
  bed_code TEXT,
  bed_desc TEXT
);
```
```{sql, connection=mydb}
CREATE TABLE IF NOT EXISTS business (
  ims_org_id TEXT PRIMARY KEY,
  business_name TEXT,
  ttl_license_beds INTEGER,
  ttl_census_beds INTEGER,
  ttl_staffed_beds INTEGER,
  bed_cluster_id INTEGER,
  FOREIGN KEY(bed_cluster_id) REFERENCES bed_type(bed_id)
);
```
```{sql, connection=mydb}
CREATE TABLE IF NOT EXISTS bed_fact (
  ims_org_id TEXT,
  bed_id INTEGER REFERENCES bed_type(bed_id),
  license_beds INTEGER,
  census_beds INTEGER,
  staffed_beds INTEGER,
  FOREIGN KEY(ims_org_id) REFERENCES business(ims_org_id),
  FOREIGN KEY(bed_id) REFERENCES bed_type(bed_id)
);
```

```{r}
bed_type <- read.csv("bed_type.csv", stringsAsFactors = FALSE)
bed_fact <- read.csv("bed_fact.csv", stringsAsFactors = FALSE)
business <- read.csv("business.csv", stringsAsFactors = FALSE)

dbWriteTable(mydb, "bed_type", bed_type, append = TRUE)
dbWriteTable(mydb, "bed_fact", bed_fact, append = TRUE)
dbWriteTable(mydb, "business", business, append = TRUE)

```
For better SQL practice, everything in the CSV files were made into SQL tables where the appropriate primary and foreign keys were applied.

## 3a
```{sql, connection=mydb}
SELECT b.business_name AS hospital_name, SUM(bf.license_beds) AS total_license_beds
FROM business b
JOIN bed_fact bf ON b.ims_org_id = bf.ims_org_id
WHERE bf.bed_id IN (4, 15)
GROUP BY b.business_name
ORDER BY total_license_beds DESC
LIMIT 10;
```

```{sql, connection=mydb}
SELECT b.business_name AS hospital_name, SUM(bf.census_beds) AS total_census_beds
FROM business b
JOIN bed_fact bf ON b.ims_org_id = bf.ims_org_id
WHERE bf.bed_id IN (4, 15)
GROUP BY b.business_name
ORDER BY total_census_beds DESC
LIMIT 10;
```

```{sql, connection=mydb}
SELECT b.business_name AS hospital_name, SUM(bf.staffed_beds) AS total_staffed_beds
FROM business b
JOIN bed_fact bf ON b.ims_org_id = bf.ims_org_id
WHERE bf.bed_id IN (4, 15)
GROUP BY b.business_name
ORDER BY total_staffed_beds DESC
LIMIT 10;
```

## 3b Interpretation of Findings


# Step 4
Query Process Steps
1. Getting Data (From, Join)
2. Row Filter (Where)
3. Grouping (Group by)
4. Group Filter (Having)
5. Return Expressions (Select)
6. Order & Paging (Order by & Limit / Offset)
## 4a
```{sql, connection=mydb}
SELECT b.business_name AS hospital_name, SUM(bf.license_beds) AS total_license_beds
FROM business b
JOIN bed_fact bf ON b.ims_org_id = bf.ims_org_id
WHERE bf.bed_id IN (4, 15)
GROUP BY b.business_name
HAVING COUNT(DISTINCT bf.bed_id) = 2
ORDER BY total_license_beds DESC
LIMIT 10;
```

```{sql, connection=mydb}
SELECT b.business_name AS hospital_name, SUM(bf.census_beds) AS total_census_beds
FROM business b
JOIN bed_fact bf ON b.ims_org_id = bf.ims_org_id
WHERE bf.bed_id IN (4, 15)
GROUP BY b.business_name
HAVING COUNT(DISTINCT bf.bed_id) = 2
ORDER BY total_census_beds DESC
LIMIT 10;
```

```{sql, connection=mydb}
SELECT b.business_name AS hospital_name, SUM(bf.staffed_beds) AS total_staffed_beds
FROM business b
JOIN bed_fact bf ON b.ims_org_id = bf.ims_org_id
WHERE bf.bed_id IN (4, 15)
GROUP BY b.business_name
HAVING COUNT(DISTINCT bf.bed_id) = 2
ORDER BY total_staffed_beds DESC
LIMIT 10;
```











